---
title: "Title of Your Report"
author: "Jiaheng Li, Heye Liu, Yan Wang, Yuxin Yang"
date: "18/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(janitor)
library(tidyverse)
library(ggplot2)
library(tinytex)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
tinytex::install_tinytex()
# Load the data dictionary and the raw data and correct the variable names
raw_data <- read_csv("gss.csv")
dataset <- raw_data %>% select(caseid, age, total_children, sex, feelings_life, marital_status, average_hours_worked, income_respondent, self_rated_health)
```

## Title of your Report

# Jiaheng Li, Heye Liu, Yan Wang, Yuxin Yang
# October 18, 2020


## Abstract

This project focuses on how people's ages and their physical health conditions could affect their life happiness based on the Canadian General Social Survey. Statistical models such as linear regressions, boxplot, are used in the analyses. The results lead to the conclusion that both age and health conditions have influences on people’s life happiness.

## Introduction

With the rapid development of modern cities, people’s quality of life seems to be improving. More convenient transportations, communication equipment, and more diverse forms of entertainment provide people with comfort in life. But it seems that people’s pace of life has become urgent. Many people live a two-point life while having to find a balance of work and family. There is very little time left for personal relaxation. This directly or indirectly causes many people’s physical health to face the threat. A lot of people are starting to ask themselves: if working hard and being busy every day, adjusting themselves to a fast living pace is for living a happier and more comfortable life, is it worth sacrificing your body health? We are also part of the busy crowd and have been thinking about this question. So, we want to find the relationship between personal health conditions and life satisfaction. Given that the health condition may be age-related, we would also add age to our analysis. Canadian General Social Survey recorded information about the basic living conditions of Canadian people including their age and self-rated health conditions, and how they feel about their current life. In the following part of the report we will introduce our selected data, the statistical models that help us visualize the relationship between the variables age and health condition as Xs’ and our Y value, the score of “feelings about life as a whole”. Eventually give a conclusion based on our analysis result, about how people’s age and health conditions may bring whether positive or negative influences to their life satisfaction. More detailed information about the variables and models will be further explained in the sections after.


## Data

  We chose to analyze the General social survey data on Family (cycle 31), 2017 from the U of T library, provided by Statistics Canada under the terms of the Data Liberation Initiative (DLI) license. The survey collected a large amount of data for each selected respondent as well as some information about each member of the respondent’s household, including age, sex, family origins, conjugal history, the respondent’s children, financial condition, labor market activity, subjective well-being and feelings about life as a whole. 
  
  The target population of this survey includes all persons 15 years of age and older in Canada. It used the redesigned GSS frame, integrating data from sources of both landline and cellular phone numbers available to Statistics Canada and the Address Register. The sample population was obtained with stratification. Provinces and many Census Metropolitan Areas (CMAs) were considered as separate stratum, and each record in the survey frame was assigned to a stratum. Then a simple random sample without replacement of records was performed in each stratum. Any household with no telephone or at least one person 15 years of age or older was excluded from the survey. Respondents were then randomly selected from eligible households and participated in the survey through telephone interview. The survey involved 20602 respondents in total and recorded 81 variables.
  
  We downloaded the raw csv dataset for STATA, cleaned and prepared it in R with the gss_cleaning file provided ahead alongside the problem set question. The gss_cleaning file converts the raw dataset into a dictionary type dataset with renamed variables. In order to analyze the effect of personal life situation on the life satisfaction score, we select the variables below:
  
+ sex: the sex of the respondent;

+ marital_status: the marital status of the respondent;

+ total_children: the total number of children the respondent has, as self-reported;

+ average_hours_worked: the average number of hours worked per week;

+ income_respondent:  the total income of respondent received in the calendar year 2016 from all sources before income taxes and deductions;

+ self_rated_health: self rated health of respondent, ranging from excellent to poor;

+ feelings_life: feelings about life as a whole using a scale of 0 to 10, where 0 means "very dissatisfied" and 10 means "very satisfied".



## Results
Firstly, we create a multiple linear model for feelings_life with self_rated_health vs ages.

***model***:

```{r echo=FALSE}
library(moderndive)
dataset$self_rated_health <- as.character(dataset$self_rated_health)
dataset$self_rated_health[dataset$self_rated_health == "Don't know"] <- 0
dataset$self_rated_health[dataset$self_rated_health == "Poor"] <- 1
dataset$self_rated_health[dataset$self_rated_health == "Fair"] <- 2
dataset$self_rated_health[dataset$self_rated_health == "Good"] <- 3
dataset$self_rated_health[dataset$self_rated_health == "Very good"] <- 4
dataset$self_rated_health[dataset$self_rated_health == "Excellent"] <- 5
dataset$age[dataset$age <= 20] <- "0-20"
dataset$age[20.1 <= dataset$age  & dataset$age<= 40] <- "20-40"
dataset$age[40.1 <= dataset$age  & dataset$age<= 60] <- "40-60"
dataset$age[60.1 <= dataset$age  & dataset$age<= 80] <- "60-80"
dataset$age[80.1 <= dataset$age] <- "80-100"

model <- lm(feelings_life ~ age + self_rated_health, data = dataset)
get_regression_table(model)
```

In addition, we make four different bar plots to reflects the impact of physical fitness on people's feeling_life at different ages. In these plots, we use use (1, 2, 3, 4, 5)  to correspond to (poor, fair, good, verygood, excellent) in the variable self_rated_health. Moreover, The age groups of these four bar plots are 0-20, 20-40, 40-60 and 60-80 respectively.


***Figure1***:

```{r echo=FALSE}
dataset$feelings_life[dataset$feelings_life == 1] <- "1-2"
dataset$feelings_life[dataset$feelings_life == 2] <- "1-2"
dataset$feelings_life[dataset$feelings_life == 3] <- "3-4"
dataset$feelings_life[dataset$feelings_life == 4] <- "3-4"
dataset$feelings_life[dataset$feelings_life == 5] <- "5-6"
dataset$feelings_life[dataset$feelings_life == 6] <- "5-6"
dataset$feelings_life[dataset$feelings_life == 7] <- "7-8"
dataset$feelings_life[dataset$feelings_life == 8] <- "7-8"
dataset$feelings_life[dataset$feelings_life == 9] <- "9-10"
dataset$feelings_life[dataset$feelings_life == 10] <- "9-10"
#show a box plot for income_respondent and feelings_life
dd <- dataset %>% filter(age == "0-20")
dd <- dd[order(dd$self_rated_health), ]
ggplot(dd, aes(self_rated_health)) + geom_bar(aes(fill = feelings_life))
```

For Figure1, we can see that in the ages of 0-20, the more healthy people are, the greater the percentage of those with a happiness index of 9 to 10. From the graph, we can clearly observe that the healthier people are at this age, the happier most of them will be in life.


***Figure2***:

```{r echo=FALSE}
dd <- dataset %>% filter(age == "20-40")
dd <- dd[order(dd$self_rated_health), ]
ggplot(dd, aes(self_rated_health)) + geom_bar(aes(fill = feelings_life))
```

For figrue2, it shows that in the age group of 20-40, There were more participants in this age group than in the 0-20 age group. And when the health status of participants at this age is good, very good or excellent, The proportion of people with a life happiness index less than or equal to 4 is higher than those with the same health status in the age group of 0-20. The reason for this may be that most people in this age group are under a lot of pressure from work, society and family. These kinds of stress may affect their level of feeling life. However, it is the same as the 0-20 age group, at this age group, a person's health status is about good, the happiness index will be high.


***Figure3***:

```{r echo=FALSE}
dd <- dataset %>% filter(age == "40-60")
dd <- dd[order(dd$self_rated_health), ]
ggplot(dd, aes(self_rated_health)) + geom_bar(aes(fill = feelings_life))
```

For figrue3, we cam make the same conclusion as the age group for 20-40,  the healthier people are at this age, the happier most of them will be in life. The proportion of people with different levels of feeling life in different health states is basically the same as that of the 20-40 age group. 


***Figure4***:

```{r echo=FALSE}
dd <- dataset %>% filter(age == "60-80")
dd <- dd[order(dd$self_rated_health), ]
ggplot(dd, aes(self_rated_health)) + geom_bar(aes(fill = feelings_life))
```

For figrue4, we also can get the conclusion that at this age group, a person's health status is about good, the happiness index will be high. There is a difference at this age group that among the people whose health status is "good", the proportion of people with a happiness index of 9-10 is higher than that of the previous three age groups. This difference may be caused by the difference in age. In the 60-80 age group, people's health is generally not as good as before, so they are not too nervous about the decline of their physical health. Moreover, most people in this age group have retired and do not have as much pressure as when they were young. So, the level of feeling life is a little higher than the first three ages.


## Discussion

Firstly, from the models section and results section, we can make a primary conclusion that People’s physical health has a greater impact on people’s level of feeling life than age has on people. When a person’s physical health is good, he will be more likely to have a high level of feeling life. On the contrary, when a person’s physical health is not good, it will be difficult for him to have a good level of feeling life. It just like an artical named Study suggests health, not wealth, determines happiness, said that "Self-described "healthy" people are 20 percent happier than average, while "unhealthy" people are 8.25 percent less happy."[1]. 

The reason for this phenomenon is that physical unhealthy or illness will cause people's lives and psychology to be greatly affected."Physical changes from a disease may affect your appearance. These changes can turn a positive self-image into a poor one. When you don't feel good about yourself, you may prefer to be alone and withdraw from friends and social activities."[2]. Moreover, physical unhealthy and illness also "can also influence your ability to work. Morning stiffness, decreased range of motion, and other physical limitations may force you to change your work activities and environment. A decreased ability to work may also lead to financial problems."[3] In addition, hysical unhealthy and illness also can cause people have more stress. "Stress can build and can shape your feelings about life. Prolonged stress can lead to frustration, anger, hopelessness, and, at times, depression."[4] Therefore, When your body was not very healthy or sick, it would have a certain impact on your life, psychology, finances, work, etc. And often these factors are the main reference conditions for people to evaluate their level of feeling life.

However, different age groups will also have a little impact on people's level of feeling life. Among these participants, the higher the age, the proportion of their level of feeling life 9-10 is relatively higher than that of other age groups, even if their health status is not very good. In our opinion, the reason why have this kind of difference is that in the 60-80 age group, people's health is generally not as good as before, so they are not too nervous about the decline of their physical health. Moreover, most people in this age group have retired and do not have as much pressure as when they were young. At the same time, Most people already have enough savings at this age. There are children to take care of them, and lovely grandchildren often come to accompany them, so the impact of unhealthy health on them will be greatly reduced by these factors. The article named Why people get happier as they get older also have the same conclusion, "Although as people move towards old age they lose things they treasure — vitality, mental sharpness and looks — they also gain what people spend their lives pursuing: happiness."[5]

Moreover, we also discover that for people aged 0-20, the proportion of people whose level of feeling life is less than 4 is smaller than that of the other three age groups. We think this is because most people in this age group have not yet started work and bear the pressure from society and family. They only need to study and do what they are interested in, and do not need to consider other. In the lives of these young people, there are not many factors that affect their level of feeling life, so the proportion of these young people whose level of feeling life is 9-10 is higher than that of the other three age groups.

In general, We conclude from this survey that age and self health status will affect people’s level of feeling life, but self health status will have a greater impact on people’s level of feeling life. This is also in line with our expectations.

# Weaknesses

There are some imperfections in the overall study which were hard to avoid. One important thing to notice is that the health condition we got from the survey was self-rated health condition, which might not be accurate real health condition. Also, people may tend to present themselves in a positive way in a society. This caused a response bias which may affect our results. There exist a lot of NA response in the survey data as well. Which makes our investigations have some blind spots.

# Next Steps


## Reference
[1] https://www.zdnet.com/article/study-suggests-health-not-wealth-determines-happiness

[2] https://my.clevelandclinic.org/health/articles/4062-chronic-illness

[3] https://my.clevelandclinic.org/health/articles/4062-chronic-illness

[4] https://my.clevelandclinic.org/health/articles/4062-chronic-illness

[5] https://medium.economist.com/why-people-get-happier-as-they-get-older-b5e412e471ed

RStudio Team (2020). RStudio: Integrated Development for R. RStudio, PBC, Boston, MA URL http://www.rstudio.com/

Yihui Xie andJ.J. Allaire and Garrett Grolemund (2018). R Markdown:The Definitive Guide. Chapman and Hall/CRC. URL https://bookdown.org/yihui/rmarkdown.







